---
title: "lab4"
output: html_document
---

```{r}

library(runjags)
library(ggplot2)
library(coda)

setwd("C:/Users/Gorge-PC/Documents")
dramadata = read.csv("KDramaData.csv", header=T)
```
question 1
```{r}
modelString <-"
model {

## likelihod
for(i in 1:N){
  y[i] ~ dnorm(mu_j[schedule[i]], invsigma2)
}

## priors

for (j in 1:J){
  mu_j[j] ~ dnorm(mu, invtau2)T(0,)
}

invsigma2 ~ dgamma(a_g, b_g)
sigma <- sqrt(pow(invsigma2, -1))

##hyperpriors
mu ~ dnorm(mu0, 1/g0^2)
invtau2 ~ dgamma(a_t, b_t)
tau <- sqrt(pow(invtau2, -1))

}
"

y = dramadata$Rating        # The y values are in the column named Rating
schedule = dramadata$Schedule      # The schedule index is in the column named Schedule
N = length(y)  # Compute the number of observations
J = length(unique(schedule)) # Compute the number of schedules/subsamples

initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}

the_data <- list("y" = y, "schedule" = schedule, "N" = N, "J" = J, 
                 "mu0" = 0.1, "g0" = 0.5, 
                 "a_t" = 1, "b_t" = 1,
                 "a_g" = 1, "b_g" = 1)

posterior <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("mu", "tau", "mu_j", "sigma"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 1, 
                      inits = initsfunction)

options(digits = 3)
summary(posterior)

plot(posterior, vars = "mu_j[1]")

```

The posterior draws of mu and mu1 through mu4 are indeed positive. MCMC diagnostics probably fit this distribution the best. The autocorrelation is near 0 after the first few draws. The traceplot varies nicely and explores the entire parameter space. 

question 2
```{r}
modelString <-"
model {

## likelihod
for(i in 1:N){
y[i] ~ dnorm(mu_j[schedule[i]], invsigma2)
}

## priors

for (j in 1:J){
mu_j[j] ~ dlnorm(mu, invtau2)
}

invsigma2 ~ dgamma(a_g, b_g)
sigma <- sqrt(pow(invsigma2, -1))

##hyperpriors
mu ~ dlnorm(mu0, 1/g0^2)
invtau2 ~ dgamma(a_t, b_t)
tau <- sqrt(pow(invtau2, -1))

}
"

y = dramadata$Rating        # The y values are in the column named Rating
schedule = dramadata$Schedule      # The schedule index is in the column named Schedule
N = length(y)  # Compute the number of observations
J = length(unique(schedule)) # Compute the number of schedules/subsamples

initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}

the_data <- list("y" = y, "schedule" = schedule, "N" = N, "J" = J, 
                 "mu0" = 0.1, "g0" = 0.5, 
                 "a_t" = 1, "b_t" = 1,
                 "a_g" = 1, "b_g" = 1)

posterior <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("mu", "tau", "mu_j", "sigma"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 1, 
                      inits = initsfunction)

options(digits = 3)
summary(posterior)

plot(posterior, vars = "mu_j[1]")
```

The posterior draws of mu and mu1 through mu4 are indeed positive. The traceplot looks better than the truncated normal distribution in that it is more varied. However, the autocorrelation doesn't drop down to 0 as quickly as it did in question one which implies it might need more thinning or a larger burn in. 

question 3
```{r}
modelString <-"
model {

## likelihod
for(i in 1:N){
y[i] ~ dnorm(mu_j[schedule[i]], invsigma2)
}

## priors

for (j in 1:J){
mu_j[j] ~ dweibull(1, mu)T(0,)
}

invsigma2 ~ dgamma(a_g, b_g)
sigma <- sqrt(pow(invsigma2, -1))

##hyperpriors
mu ~ dlnorm(mu0, 1/g0^2)

}
"

y = dramadata$Rating        # The y values are in the column named Rating
schedule = dramadata$Schedule      # The schedule index is in the column named Schedule
N = length(y)  # Compute the number of observations
J = length(unique(schedule)) # Compute the number of schedules/subsamples

initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}

the_data <- list("y" = y, "schedule" = schedule, "N" = N, "J" = J, 
                 "mu0" = 0.1, "g0" = 0.5, 
                 "a_t" = 1, "b_t" = 1,
                 "a_g" = 1, "b_g" = 1)

posterior <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("mu", "tau", "mu_j", "sigma"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 1, 
                      inits = initsfunction)

options(digits = 3)
summary(posterior)

plot(posterior, vars = "mu_j[1]")

```

The posterior draws of mu and mu1 through mu4 are indeed positive. This is using a weibull distribution prior with a log normal hyperprior. The autocorrelation looks very similar to question 2. It might be slightly better, but not by much. The traceplot is also varied. It would likely also benefit from a larger burn in and more thinning. 