---
title: "lab5"
output: html_document
---

```{r}
library(runjags)
library(ggplot2)
library(coda)
library(gridExtra)
library(tidyverse)
library(ggridges)

CEData = read.csv("CEsample.csv",
                  header = T, sep = ",")


```
1.
```{r}
## write the model
modelString <-"
model {
## sampling
for (i in 1:N){
y[i] ~ dnorm(beta0 + beta1*x[i], invsigma2)
}
## priors
beta1 <- (mu2 - mu1)/(x2 - x1)
beta0 <- mu1 - x1*(mu2 - mu1)/(x2 - x1)
mu1 ~ dnorm(m1, g1)
mu2 ~ dnorm(m2, g2)
invsigma2 ~ dgamma(a, b)
sigma <- sqrt(pow(invsigma2, -1))
}
"
y <- as.vector(CEData$log_TotalExp)
x <- as.vector(CEData$log_TotalIncome)
N <- length(y)
the_data <- list("y" = y, "x" = x, "N" = N,
                 "m1" = 8, "g1" = 0.25,
                 "m2" = 11, "g2" = 0.25,
                 "x1" = 10, "x2" = 12,
                 "a" = 1, "b" = 1)

posterior <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("beta0", "beta1", "sigma"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000)

summary(posterior)
plot(posterior)
```
2.
This method overall looks very solid. Autocorrelations of the parameters are generally low and trace plots look like they do a good job of exploring the full parameter space. 

3. 
B0 in this case is the log expenditure for a log income of 0. B1 would be the change in log expenditure for an increase in log income of 1. 

4.
```{r}
post <- as.mcmc(posterior)
post <- as.data.frame(post)

one_predicted <- function(x){
lp <- post[ , "beta0"] + x * post[ , "beta1"]
y <- rnorm(5000, lp, post[, "sigma"])
data.frame(Value = paste("Size =", x),
Predicted_logExp = y)
}
df <- map_df(c(1, 5, 7, 9), one_predicted)

ggplot(df, aes(x = Predicted_logExp, y = Value)) +
geom_density_ridges() +
theme_grey(base_size = 18, base_family = "")

df %>% group_by(Value) %>%
summarize(P05 = quantile(Predicted_logExp, 0.05),
          P50 = median(Predicted_logExp),
          P95 = quantile(Predicted_logExp, 0.95)
)
```

A log income level of \$5 is estimated to have a log expenditure of about \$6.42. There's a 90% certainty that the log expenditure falls between \$5.23 and \$7.63. 

