---
title: "Midterm Take Home"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(TeachBayes)
```
1.

a.
```{r}
###2005
S = 1000
prior = rbeta(S, 1, 1)

alpha = 1
beta = 0

mu = 264/(264+1496)
n = 264 + 1496

alpha_n = alpha+n/2
beta_n = beta+1/2*sum((prior-mu)^2)

set.seed(123)
p2005_post = rgamma(S,shape=alpha_n,rate=beta_n)
df = as.data.frame(p2005_post)

ggplot(df, aes(p2005_post)) + 
  geom_density(size = 2, color = "blue") + 
  labs(title = "Posterior draws of invsigma2_post") +
  theme_grey(base_size = 30, base_family = "") 

###2015

mu2 = 437/(437+1495)
n2 = 437 + 1495

alpha_n2 = alpha+n2/2
beta_n2 = beta+1/2*sum((prior-mu2)^2)

set.seed(123)
S = 1000
p2015_post = rgamma(S,shape=alpha_n2,rate=beta_n2)
df = as.data.frame(p2015_post)

ggplot(df, aes(p2015_post)) + 
  geom_density(size = 2, color = "blue") + 
  labs(title = "Posterior draws of invsigma2_post") +
  theme_grey(base_size = 30, base_family = "") 

```
b.
```{r}

diffpost <- setdiff(p2015_post, p2005_post)
sum(diffpost > 0 | diffpost < 0)/S
```
Yes, the proportions did change. The probability that 

c.
```{r}
p2005 = 264/(264+1496)
odds2005 = p2005/(1-p2005)
odds2005

p2015 = 437/(437+1495)
odds2015 = p2015/(1-p2015)
odds2015
```
Yes. The odds did change. 

d.
There is an assumption made that a science and non-science major is mutually exclusive. People who decided to double major, one in science and one not, would be counted twice when calculating the proportion because they show up once under science and once under non-science. This can deflate the proportion of science majors in either year.



2.

a.
```{r}
data <- c(9, 7.5, 7, 8, 5, 6.5, 8.5, 7, 9, 7, 5.5, 6, 8.5, 7.5)

mu_0 = 8
sigma_0 = 1
phi_0 = 1/sigma_0^2

ybar = sum(data)/16
phi = 1/(1.5^2)
n = 16

mu_n = (phi_0*mu_0+n*ybar*phi)/(phi_0+n*phi)
sd_n = sqrt(1/(phi_0+n*phi))

set.seed(123)
S = 1000
mu_post = rnorm(S, mean = mu_n, sd = sd_n)
df = as.data.frame(mu_post)

ggplot(df, aes(mu_post)) + 
  geom_density(size = 2, color = "blue") + 
  labs(title = "Posterior draws of mu") +
  theme_grey(base_size = 30, base_family = "") 
```

b.
```{r}
qnorm(c(0.05, 0.95), mean = mu_n, sd = sd_n)
```

c.
```{r}
S = 1000
sd_sim = (1/phi)+(1/phi_0)

set.seed(123)
pred_y_sim <- rnorm(S, mu_n, sd_sim)
df2 = as.data.frame(pred_y_sim)

ggplot(df2, aes(pred_y_sim)) + 
  geom_density(size = 2, color = "blue") + 
  labs(title = "Predictive Distribution") +
  theme_grey(base_size = 30, base_family = "") 

qnorm(c(0.05, 0.95), mean = mu_n, sd = sd_sim)
```

3.

a.
```{r}
data <- c(9, 7.5, 7, 8, 5, 6.5, 8.5, 7, 9, 7, 5.5, 6, 8.5, 7.5)

mu_0 = 8
sigma_m_0 = 0.1
phi_m_0 = 1/sigma_m_0^2

ybar = sum(data)/16
phi = 1/(1.5^2)
n = 16

mu_m_n = (phi_m_0*mu_0+n*ybar*phi)/(phi_m_0+n*phi)
sd_m_n = sqrt(1/(phi_m_0+n*phi))

set.seed(123)
S = 1000
mu_m_post = rnorm(S, mean = mu_m_n, sd = sd_m_n)
df3 = as.data.frame(mu_m_post)

ggplot(df3, aes(mu_m_post)) + 
  geom_density(size = 2, color = "blue") + 
  labs(title = "Posterior draws of mu") +
  theme_grey(base_size = 30, base_family = "") 
```

b.
```{r}
qnorm(c(0.05, 0.95), mean = mu_m_n, sd = sd_m_n)

```

c.

Yes, the credible interval is very sensitive to the choice of prior. The range of the 90% credible interval using Mary's prior, is only roughly 0.25. Whereas, with John's prior, it is roughly 1. Mary's interval is also not even entirely within John's interval. There isn't huge overlap between the two ranges. Thus, the credible interval is very different.
